---
title: "visualization_finance_example"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stringr)
```
## data visualization practice
```{r trim and multi fun fxn , include=FALSE}
trim <- function (x) gsub("^\\s+|\\s+$", "", x)   # function to trim spaces of columns
multi.fun <- function(x) {cbind(freq = table(x),
                                percentage = prop.table(table(x))*100)} 
## loading documents
inpatientCharges <- read_csv("~/Documents/inpatientCharges.csv", 
                             col_types = 
                               cols(`Average Covered Charges` = col_number(), 
        `Average Medicare Payments` = col_number(), 
        `Average Total Payments` = col_number()))
colnames(inpatientCharges) = 
  tolower(stringr::str_replace_all(colnames(inpatientCharges), " ", "_"))
inpatientCharges$provider_id <- as.character(inpatientCharges$provider_id)
```
```{r forgot, also replace missings fxn and reorder fxn }
inpatientCharges$provider_zip_code <- as.character(inpatientCharges$provider_zip_code)
replace_missings <- function(x, replacement) {
  is_miss <- is.na(x)
  x[is_miss] <- replacement

  message(sum(is_miss), " missings replaced by the value ", replacement)
  x
}
reorder_size <- function(x) {
        factor(x, levels = names(sort(table(x), decreasing = TRUE)))
}
# oduble checking for missing data
sapply(inpatientCharges, function(x) sum(length(which(is.na(x)))))
```
```{r using psych package }
psych::describe(inpatientCharges[num])
```

```{r}
head(inpatientCharges, 2) # looking at head of inpatientCharges
```
```{r creation of hospital state max count df }
hospital_state_max_count_df <- inpatientCharges %>% 
  separate(hospital_referral_region_description, c("hospital_state", "hospital_city"), sep = " - ") %>% 
  group_by(hospital_state) %>% 
  summarize(max_covered = max(average_covered_charges), count = n(), max_disch = max(total_discharges))
```
```{r tidy df }
inpatientCharges.tidy <-
  inpatientCharges %>% 
  separate(hospital_referral_region_description, c("hospital_state", "hospital_city"), sep = " - ", remove = F) %>% 
  separate(drg_definition, c("drg_code", "drg_description"), sep = " - ", remove = F)

```
```{r}
library(noncensus)
data(states) # data from noncensus package
hospital_states <- states %>% select(-capital)
colnames(hospital_states) <- c("hospital_state", "hospital_state_name","hospital_state_region", "hospital_state_division", "hospital_state_area","hospital_state_population")
```
```{r hospital states df creation, include=FALSE, message=FALSE}
hospital_states <-
  hospital_states %>% mutate_at(vars(hospital_state_area, hospital_state_population), as.numeric) %>% mutate(hospital_pop_area = round(hospital_state_population/hospital_state_area, 4))
inpatientCharges.tidy1 <- inpatientCharges.tidy %>% left_join(hospital_states)
```
```{r describing by patient region}
psych::describeBy(group = inpatientCharges.tidy1$hospital_state_region, x=inpatientCharges.tidy1$average_medicare_payments)
```
```{r, message= FALSE, include=FALSE}
b <- inpatientCharges  %>% distinct(drg_definition) %>% mutate(drg_definition_last = str_sub(drg_definition, start=-12)) %>% separate(drg_definition_last, c("drg_a", "drg_b"), sep= " W", remove=T) %>% mutate(drg_W = paste0("W", drg_b))
newcat <- b %>% select(drg_W, drg_definition)
```
```{r}
inpatientCharges.tidy.v <- newcat %>% right_join(inpatientCharges.tidy1,by = "drg_definition")
```
```{r}
library(ggsci) # ggsci because of the futurama color scheme
theme_set(theme_bw()) # themeset bw
library(ggplot2)
library(gridExtra)
inpatientCharges.tidy.v  %>% 
  filter(provider_street_address=="550 FIRST AVENUE") %>%
  ggplot(aes(x = average_medicare_payments/total_discharges, y = average_covered_charges/total_discharges)) +
  geom_smooth() +
   geom_point(alpha=0.9, pch=21, col="purple", size=2) + 
  scale_color_futurama() + scale_x_log10() + scale_y_log10() +
  ggtitle("NYU LANGONE 550 FIRST AVE")
```
_Extremely busy plot/ not helpful_
```{r hospital statedf}
head(hospital_state_max_count_df,2) #looking at hospital_state_max_count_df
```
```{r creating large color palette }
library(randomcoloR) # loading randomcolorR
n <- 51
palette_51 <- distinctColorPalette(n) # creating 51 random color palettes
```

```{r}
ggplot(hospital_state_max_count_df, aes(x = max_covered, y = max_disch, size = count, col=hospital_state)) + geom_point(pch=21) +
  geom_point(alpha=0.2) + ggtitle("Inpatient Maximum Covered Payments and Maximum Discharges by State") + theme_bw() +
  scale_color_manual(values=palette_51)+ guides(col = guide_legend(nrow = 10
                                                                   ), size=guide_legend(nrow=1)) +
  labs(x="Maximum Discharges by State", y="Maximum Covered Inpatient Payments")
```
# looking at MCC / CC
```{r}
colnames(inpatientCharges.tidy.v)
inpatient_region_total_discharges_summary <- psych::describeBy(group= inpatientCharges.tidy.v$hospital_state_region, x=inpatientCharges.tidy.v$total_discharges)
inpatient_region_average_medicare_payments_summary <- psych::describeBy(group= inpatientCharges.tidy.v$hospital_state_region, x=inpatientCharges.tidy.v$average_medicare_payments)
inpatient_region_average_medicare_payments_summary <- psych::describeBy(group= inpatientCharges.tidy.v$hospital_state_region, x=inpatientCharges.tidy.v$average_medicare_payments)
```

```{r changing name to df.t to make it easier }
df.t <- inpatientCharges.tidy.v
df.t %>% filter(provider_street_address=="550 FIRST AVENUE") %>%
  filter(drg_W == "W MCC" | drg_W=="W CC") %>%
  ggplot(aes(x=drg_W, y=average_covered_charges, fill=drg_W)) + geom_boxplot()+ scale_fill_brewer(palette = "Dark2") +
  labs(title="NYU 550 First Avenue",
subtitle="Inpatient Covered Charges by if procedure had complications (CC), major complications (MCC)", x="Procedure Complication")# clearly skewed
```
_If procedure had MCC vs W CC, then higher covered charges_

```{r}
part1 <- inpatientCharges.tidy.v%>% filter(hospital_state=="NY")  %>%
  select(hospital_state, drg_code, provider_name)
inpatientCharges.tidy.v %>% filter(hospital_state=="NY") %>% group_by(provider_name) %>% summarize(tot_discharges=mean(total_discharges), average_tot_payments=mean(average_total_payments)) %>% left_join(part1, by="provider_name") %>% mutate(`NYU HOSPITAL`=if_else(provider_name=="NYU HOSPITALS CENTER", "YES", "NO")) %>% ggplot(aes(tot_discharges, y=average_tot_payments, col=`NYU HOSPITAL`)) +
  geom_count(alpha=0.2) +geom_count(alpha=0.92, pch=21) +
    scale_color_manual(values=c("#9DE0AD","#8C5881")) +
  labs(title="Average Total Payments vs Average Total Discharges by Hospital in NY",
       subtitle="Size=Counts", caption="source: data.gov",
       y="Average Total Payments By Hospital",
       x="Average Total Discharges By Hospital")

```

```{r}
inpatientCharges.tidy.v %>% filter(hospital_state=="NY") %>% group_by(drg_code) %>% summarize(tot_discharges=mean(total_discharges), average_tot_payments=mean(average_total_payments)) %>% ggplot(aes(x = tot_discharges, y = average_tot_payments)) +
  geom_point(alpha=0.5) + 
  scale_x_log10() +
  labs(title="Average Total Discharges vs Average Total Payments for DRG procedure in NY",
       subtitle="Log X scale") 


```
```{r creation of just subset with city as manhattan }
df.manhattan.inpatient.expenses<- df.t %>% filter(hospital_referral_region_description== "NY - Manhattan")
df.manhattan.inpatient.expenses %>% head(2)
multi.fun(df.manhattan.inpatient.expenses$provider_city)
```

```{r}
numeric.manhattan <- sapply(df.manhattan.inpatient.expenses, is.numeric)
df.manhat.cat <- df.manhattan.inpatient.expenses[!numeric.manhattan]
df.manhat.cat %>% select(provider_id, provider_name) %>% head(2)
```
```{r}
df.manhattan.inpatient.expenses$NYU=factor(ifelse(df.manhattan.inpatient.expenses$provider_name=="NYU HOSPITALS CENTER","NYU","NOT NYU"))

df.manhattan.inpatient.expenses$withcomplicationsorwithout <- gsub("W CC/MCC", "W COMPLICATIONS", df.manhattan.inpatient.expenses$drg_W)
df.manhattan.inpatient.expenses$withcomplicationsorwithout <- gsub("W MCC", "W COMPLICATIONS", df.manhattan.inpatient.expenses$withcomplicationsorwithout)
df.manhattan.inpatient.expenses$withcomplicationsorwithout <- gsub("W CC", "W COMPLICATIONS", df.manhattan.inpatient.expenses$withcomplicationsorwithout)
df.manhattan.inpatient.expenses$withcomplicationsorwithout <- gsub("W/O CC/MCC", "W/O COMPLICATIONS", df.manhattan.inpatient.expenses$withcomplicationsorwithout)
df.manhattan.inpatient.expenses$withcomplicationsorwithout <- gsub("W/O MCC", "W/O COMPLICATIONS", df.manhattan.inpatient.expenses$withcomplicationsorwithout)
df.manhattan.inpatient.expenses$withcomplicationsorwithout <- gsub("WNA", "DOES NOT APPLY TO PROCEDURE", df.manhattan.inpatient.expenses$withcomplicationsorwithout)
multi.fun(df.manhattan.inpatient.expenses$drg_W)
multi.fun(df.manhattan.inpatient.expenses$withcomplicationsorwithout)
```

How does NYU compare to other hospitals inpatient expenses?
```{r}
df.manhattan.inpatient.expenses %>% mutate(avg_med_pay_per_disch = round(average_medicare_payments/total_discharges,2)) %>% ggplot(aes(x=reorder(provider_name, avg_med_pay_per_disch, FUN=median), y=avg_med_pay_per_disch)) + geom_boxplot(varwidth = TRUE, aes(fill=NYU)) + coord_flip() +
    scale_alpha_manual(values=c(0.2,0.1)) +
    scale_fill_manual(values=c("#9DE0AD","#8C5881")) +
  labs(x="Hospital", y="Average Medicare Payments / Total Discharges", 
       title = "Boxplots",
       subtitle=
         "ordered by median medicare payments per discharge") +
   theme(legend.position = "none")
```
```{r}
df.manhattan.inpatient.expenses %>% group_by(provider_name) %>% summarize(count=n(), average_disch = round(mean(total_discharges))) %>% head(2)
# %>% left_join(df.manhattan.inpatient.expenses)
```

```{r what about nyu total payments in comparison to other NYC region hospitals }
df.manhattan.inpatient.expenses %>% ggplot(aes(x=reorder(provider_name, average_total_payments, FUN=median), y=average_total_payments)) + geom_boxplot(varwidth = T, aes(fill=NYU)) + coord_flip() +
    scale_alpha_manual(values=c(0.2,0.1)) +
    scale_fill_manual(values=c("#9DE0AD","#8C5881")) +
  labs(x="Hospital", y="Average Total Payments", 
       subtitle=
         "Boxplots ordered by median of Average Total Payments") +
  theme(legend.position = "none")
```
```{r}
describeBy_total_discarges <- psych::describeBy(df.manhattan.inpatient.expenses$total_discharges, df.manhattan.inpatient.expenses$provider_name)
 NYU.TOTAL.DISCHARGES <- data.frame(describeBy_total_discarges$`NYU HOSPITALS CENTER`)
 rownames(NYU.TOTAL.DISCHARGES) <- c("total.discharges.NYU")
 numeric <- sapply(df.t, is.numeric)
df.manhattan.inpatient.expenses %>% select(total_discharges, average_covered_charges, average_medicare_payments, average_total_payments, provider_name)
```
```{r dotplot }
manhattan.hospital.procedures <- df.manhattan.inpatient.expenses %>% group_by(provider_name) %>% tally()
df.manhattan.inpatient.expenses %>% select(NYU, provider_name) %>% distinct() %>% right_join(manhattan.hospital.procedures, by = "provider_name") %>%
  ggplot(aes(reorder(provider_name, n), n, col=NYU)) + geom_point() + coord_flip() +
    scale_color_manual(values=c("#B7D968","#8C5881")) +
  labs(subtitle="DRG procedures performed at Manhattan Hospitals", y="Number of distinct DRG Procedures performed", x="Hospital") +
  theme(legend.position = 
          "none")
```

```{r}
# df.manhattan.inpatient.expenses %>% ggplot(aes(x=reorder(provider_name,total_discharges, FUN=median), y=total_discharges)) + geom_boxplot(varwidth = T, aes(fill=NYU)) + coord_flip() +
#     scale_alpha_manual(values=c(0.2,0.1)) +
#     scale_fill_manual(values=c("#9DE0AD","#8C5881")) +
#   labs(x="Hospital", y="Average Total Payments", 
#        subtitle=
#          "Boxplots ordered by max total discharges")
```

```{r creation of manhattan plot }
manhattan_plot <- ggplot(df.manhattan.inpatient.expenses, aes(x = average_medicare_payments, y = average_total_payments))
# using provider city as color
manhattan_plot+ geom_jitter(aes(color = provider_city), pch=21) + scale_color_futurama()

df.550.first.ave <- df.manhattan.inpatient.expenses %>% filter(provider_street_address ==
                                             "550 FIRST AVENUE")

df.550.first.ave %>%
  group_by(withcomplicationsorwithout) %>% 
  summarize(max_covered = max(average_covered_charges), count = n(), max_disch = max(total_discharges))
```
```{r}
sapply(df.550.first.ave[,c("average_medicare_payments","total_discharges","average_covered_charges","average_total_payments")], summary)
```

```{r}
drg_W_df.550.first.ave <- psych::describeBy(df.550.first.ave$average_total_payments, group=df.550.first.ave$drg_W)
drg_W_df.550.first.ave$WNA

```

```{r looking at different NY-MANHATTAN hospitals and max covered charges}
df.manhattan.inpatient.expenses %>% 
  group_by(provider_name) %>% 
  summarize(max_covered_charges = max(average_covered_charges), max_medicare_payments= max(average_medicare_payments),count = n(), max_discharges = max(total_discharges), max_total_payments=max(average_total_payments))
```
```{r}
x <- sapply(df.550.first.ave, is.numeric)
colnames(df.550.first.ave[x])
```
```{r 550 first ave procedures with major complications ("W MCC ")}
df.550.first.ave %>% filter(drg_W=="W MCC") %>% ggplot(aes(average_medicare_payments/total_discharges, y=average_total_payments/total_discharges)) + geom_jitter(aes(col=drg_code, size=average_covered_charges), alpha=0.2) + geom_jitter(aes(col=drg_code, size=average_covered_charges), alpha=0.9,pch=21) + labs(subtitle="NYU LANGONE 550 FIRST AVE", title="procedures W MCC") + scale_color_manual(values=palette_51)
```
```{r creation of color palettes }
wes_palettes <- list(
  BottleRocket1 = c("#A42820", "#5F5647", "#9B110E", "#3F5151", "#4E2A1E", "#550307", "#0C1707"),
  BottleRocket2 = c("#FAD510", "#CB2314", "#273046", "#354823", "#1E1E1E"),
  Rushmore1 = c("#E1BD6D", "#EABE94", "#0B775E", "#35274A" ,"#F2300F"),
  Royal1 = c("#899DA4", "#C93312", "#FAEFD1", "#DC863B"),
  Royal2 = c("#9A8822", "#F5CDB4", "#F8AFA8", "#FDDDA0", "#74A089"),
  Zissou1 = c("#3B9AB2", "#78B7C5", "#EBCC2A", "#E1AF00", "#F21A00"),
  Darjeeling1 = c("#FF0000", "#00A08A", "#F2AD00", "#F98400", "#5BBCD6"),
  Darjeeling2 = c("#ECCBAE", "#046C9A", "#D69C4E", "#ABDDDE", "#000000"),
  Chevalier1 = c("#446455", "#FDD262", "#D3DDDC", "#C7B19C"),
  FantasticFox1 = c("#DD8D29", "#E2D200", "#46ACC8", "#E58601", "#B40F20"),
  Moonrise1 = c("#F3DF6C", "#CEAB07", "#D5D5D3", "#24281A"),
  Moonrise2 = c("#798E87", "#C27D38", "#CCC591", "#29211F"),
  Moonrise3 = c("#85D4E3", "#F4B5BD", "#9C964A", "#CDC08C", "#FAD77B"),
  Cavalcanti1 = c("#D8B70A", "#02401B", "#A2A475", "#81A88D", "#972D15"),
  GrandBudapest1 = c("#F1BB7B", "#FD6467", "#5B1A18", "#D67236"),
  GrandBudapest2 = c("#E6A0C4", "#C6CDF7", "#D8A499", "#7294D4")
)
gbp =c("#733080","#F4BAC8","#A40607","#7288B9","#F0C595","#733080")
nyu_cols_maybe <- c(gbp, wes_palettes$FantasticFox1, wes_palettes$GrandBudapest1, wes_palettes$GrandBudapest2)

```
```{r}
ggplot(df.t, aes(x = average_total_payments, y = average_medicare_payments)) +
geom_point(alpha=0.01) + ggtitle("Average Medicare Payments vs Average Total Payments for all US") + geom_smooth(size=0.9)

```

```{r}
hospital_states$hospital_state <- as.character(hospital_states$hospital_state)
state_averages_df <-
  df.t %>% 
  group_by(hospital_state) %>% 
  summarize(state_mean_covered_charges = mean(average_covered_charges), state_mean_total_discharges = mean(total_discharges),state_average_total_payments =
mean(average_total_payments), state_average_medicare_payments = mean(average_medicare_payments)

) %>% left_join(hospital_states, by="hospital_state") %>% mutate(state_medicare_payments_over_total_payments=state_average_medicare_payments/state_average_total_payments)

```

```{r dataviz example}
ggplot(state_averages_df, aes(x = state_mean_covered_charges, y = state_average_total_payments,color = hospital_state_region)) +
  geom_point(aes(size=hospital_pop_area)) +
  geom_smooth(method="loess") + scale_color_futurama()
```
```{r}
ggplot(state_averages_df, aes(x = state_mean_covered_charges, y = state_average_total_payments,color = hospital_state_region)) +
  # geom_point() +
  geom_smooth() + scale_color_futurama()
```

```{r}
state_averages_df %>% ggplot(aes(hospital_state_population, state_average_total_payments, col=hospital_state_region)) + geom_jitter(alpha=0.72) + scale_color_manual(values=palette_51) 
```
```{r}
state_averages_df %>% ggplot(aes(x = 1000*state_medicare_payments_over_total_payments/sum(state_medicare_payments_over_total_payments), y = hospital_pop_area, col=hospital_state_region)) + geom_point(alpha=0.4) + scale_color_futurama() +
  labs(x="medicare payments/sum", subtitle="Hospital Population Area vs Total Medicare Payments/Sum of Total Medicare Payments")
```
```{r}
df.manhattan.inpatient.expenses %>% group_by(withcomplicationsorwithout, NYU) %>% 
  summarise(mean_covered_charge_per_disch = mean(average_covered_charges/total_discharges))
```

```{r}
df.manhattan.inpatient.expenses %>% group_by(withcomplicationsorwithout, NYU) %>% 
  summarise(mean_covered_charge_per_disch = mean(average_covered_charges/total_discharges)) %>%
  ggplot(aes(factor(withcomplicationsorwithout), mean_covered_charge_per_disch, fill=NYU)) + geom_col(position="dodge") +
  scale_fill_manual(name="", values=c("#9DE0AD","#8C5881")) + coord_flip() +
  labs(title="Covered Charges for Procedures NYU vs other Manhattan region hospitals", y="Average Covered Charge Per Discharge", caption="source: data.gov", x="Procedures: Without CC/MCC, With CC/MCC, N/A")
```
```{r NYU vs not NYU for drg_coes and average total payments }
df.manhattan.inpatient.expenses %>% group_by(drg_code, NYU) %>% 
  summarise(avg_total_payments_per_disch = mean(average_total_payments/total_discharges)) %>% head(10)
```

```{r}
df.manhattan.inpatient.expenses %>% group_by(drg_code, NYU) %>% 
  summarise(avg_total_payments_per_disch = 
              mean(average_total_payments/total_discharges)) %>%
  ggplot(aes(factor(drg_code), avg_total_payments_per_disch, fill=NYU)) + 
  geom_col(position="dodge") +
  scale_fill_manual(name="", 
                    values=c("#9DE0AD","#8C5881")) + coord_flip() +
  labs(subtitle=
         "Average total payments per discharge for DRG Procedures NYU vs other Manhattan region hospitals", y="Average total payments per discharge",
       caption="source: data.gov", x="DRG Codes") + theme(axis.text.y = element_text(size=6))
```

```{r summary of df.550.first.ave }
summary(df.550.first.ave$average_covered_charges)
```

```{r}
nyu_color_scale <- scale_color_manual(values=nyu_cols_maybe)
# 1st ave plot
firstave_plot <- ggplot(df.550.first.ave, aes(x = average_covered_charges/total_discharges, y = average_medicare_payments/total_discharges))

dia_plot <- firstave_plot + geom_point(alpha=0.2, size=3) + ggtitle("NYU HOSPITAL- 550 FIRST AVE")

dia_plot + geom_smooth(se = FALSE, col="purple")

dia_plot + geom_smooth(aes(col = withcomplicationsorwithout), se = FALSE) +nyu_color_scale 
```
```{r}
df.550.first.ave%>%
ggplot( aes(x = average_covered_charges/total_discharges, y = average_medicare_payments/total_discharges)) + geom_point(alpha=0.92,
  aes(col=withcomplicationsorwithout, size=total_discharges), pch=21)
```
```{r}
summary(df.550.first.ave$average_covered_charges/df.550.first.ave$total_discharges)
summary(df.550.first.ave$average_total_payments)
summary(df.550.first.ave$average_medicare_payments)
```

```{r}
library(ggalt)
payments_select <- df.550.first.ave[
   df.550.first.ave$average_covered_charges> 190000,]
   # & 
   #                         df.550.first.ave$average_medicare_payments > 35000, ]

ggplot(df.550.first.ave, aes(x = average_total_payments, y = average_medicare_payments)) + geom_point(alpha=0.2, size=1.3,
  aes(col=withcomplicationsorwithout)) +
  geom_point(alpha=0.7, size=1.3, pch=21,
  aes(col=withcomplicationsorwithout)) +
      # , size=total_discharges)) +   # draw points
  geom_smooth(method="loess", se=F, col="purple", size=.75) +
  # xlim(c(0, 0.1)) + 
  # ylim(c(0, 500000)) +   # draw smoothing line
  geom_encircle(aes(x=average_total_payments, y = average_medicare_payments), 
                data=payments_select, 
                color="red", 
                size=1.2, 
                expand=0.02) +   # encircle
  labs(title="Scatterplot of average medicare payments vs average total payments", 
       subtitle="Encircle of procedures where average covered charges exceed $190000", 
       caption="Source: data.gov") + scale_color_manual(values=c("purple","darkred", "black"))
```


```{r}
  ggplot(df.550.first.ave,aes(x=withcomplicationsorwithout, y=average_total_payments, fill=withcomplicationsorwithout)) + geom_boxplot(varwidth = T)+ scale_fill_manual(name="Procedure Complication", values = palette_51) +
  labs(title="NYU 550 First Avenue",
subtitle="Average Total Payments", x="Procedure Complication") +
  coord_flip()
```
```{r }
df.550.first.ave %>% ggplot(aes(average_covered_charges, average_total_payments)) + geom_point(col="purple", size=2, alpha=0.25) +
  labs(title="Covered Charges vs Total Payments at NYU 550 First Ave")
firstaveModel <- lm(average_total_payments ~ average_covered_charges, data = df.550.first.ave)
```
using base package plot
```{r}
df.550.first.ave$withcomplicationsorwithout <- as.factor(df.550.first.ave$withcomplicationsorwithout)
plot(df.550.first.ave$average_covered_charges, df.550.first.ave$average_total_payments, col = df.550.first.ave$withcomplicationsorwithout)
abline(firstaveModel, lty = 2)
```
```{r}
# install.packages("CerioliOutlierDetection")

head(df.550.first.ave[14:17])
```
```{r using Hmisc to describe numeric cols in the 550 first ave df }
Hmisc::describe(df.550.first.ave[14:17])
```
```{r}
library(gmodels)
CrossTable(df.550.first.ave$withcomplicationsorwithout)
str(df.manhattan.inpatient.expenses)
CrossTable(df.manhattan.inpatient.expenses$provider_city, df.manhattan.inpatient.expenses$withcomplicationsorwithout, prop.r = TRUE, 
           prop.c = FALSE, prop.t = FALSE, prop.chisq = FALSE)
```
We can see that the proportion of procedures with complications or without tends to stay relatively the same between all provider cities

```{r using psych package}
numerical_summary_df.550.first.ave <- as.data.frame(psych::describe(df.550.first.ave[14:17]))
 numerical_summary_df.550.first.ave$num.vars <- rownames(numerical_summary_df.550.first.ave)
 numerical_summary_df.550.first.ave # now a dataframe
```
```{r creating indexes of outliers to investigate later }
index_highdischarges <- which(df.550.first.ave$total_discharges > 680) # max is 689
# definitely an outlier as can be seen below
plot(df.550.first.ave$total_discharges)
plot(df.550.first.ave$total_discharges,df.550.first.ave$average_covered_charges)
plot(df.550.first.ave$average_covered_charges)
plot(df.550.first.ave$average_total_payments)
plot(df.550.first.ave$average_medicare_payments)
```

```{r}

n_breaks <- sqrt(nrow(df.550.first.ave))
hist(df.550.first.ave$total_discharges, xlab="Total Discharges", main="Total Discharges at NYU 550 First Ave", breaks=11)
hist(df.550.first.ave$average_covered_charges,breaks=11, xlab="Average Covered Charges", main="Avg Covered Charges at NYU 550 First Ave")
hist(df.550.first.ave$average_total_payments,breaks=11, xlab="Average Total Payments", main="Avg Covered Charges at NYU 550 First Ave")
hist(df.550.first.ave$average_medicare_payments,breaks=11, xlab="Average Medicare Payments", main="Average Medicare Payments at NYU 550 First Ave")
```

```{r}

```


From R cookbook:
```{r herlper functions}
## Gives count, mean, standard deviation, standard error of the mean, and confidence interval (default 95%).
##   data: a data frame.
##   measurevar: the name of a column that contains the variable to be summariezed
##   groupvars: a vector containing names of columns that contain grouping variables
##   na.rm: a boolean that indicates whether to ignore NA's
##   conf.interval: the percent range of the confidence interval (default is 95%)
summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
    library(plyr)

    # New version of length which can handle NA's: if na.rm==T, don't count them
    length2 <- function (x, na.rm=FALSE) {
        if (na.rm) sum(!is.na(x))
        else       length(x)
    }

    # This does the summary. For each group's data frame, return a vector with
    # N, mean, and sd
    datac <- ddply(data, groupvars, .drop=.drop,
      .fun = function(xx, col) {
        c(N    = length2(xx[[col]], na.rm=na.rm),
          mean = mean   (xx[[col]], na.rm=na.rm),
          sd   = sd     (xx[[col]], na.rm=na.rm)
        )
      },
      measurevar
    )

    # Rename the "mean" column    
    datac <- rename(datac, c("mean" = measurevar))

    datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean

    # Confidence interval multiplier for standard error
    # Calculate t-statistic for confidence interval: 
    # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
    ciMult <- qt(conf.interval/2 + .5, datac$N-1)
    datac$ci <- datac$se * ciMult

    return(datac)
}
## Norms the data within specified groups in a data frame; it normalizes each
## subject (identified by idvar) so that they have the same mean, within each group
## specified by betweenvars.
##   data: a data frame.
##   idvar: the name of a column that identifies each subject (or matched subjects)
##   measurevar: the name of a column that contains the variable to be summariezed
##   betweenvars: a vector containing names of columns that are between-subjects variables
##   na.rm: a boolean that indicates whether to ignore NA's
normDataWithin <- function(data=NULL, idvar, measurevar, betweenvars=NULL,
                           na.rm=FALSE, .drop=TRUE) {
    library(plyr)

    # Measure var on left, idvar + between vars on right of formula.
    data.subjMean <- ddply(data, c(idvar, betweenvars), .drop=.drop,
     .fun = function(xx, col, na.rm) {
        c(subjMean = mean(xx[,col], na.rm=na.rm))
      },
      measurevar,
      na.rm
    )

    # Put the subject means with original data
    data <- merge(data, data.subjMean)

    # Get the normalized data in a new column
    measureNormedVar <- paste(measurevar, "_norm", sep="")
    data[,measureNormedVar] <- data[,measurevar] - data[,"subjMean"] +
                               mean(data[,measurevar], na.rm=na.rm)

    # Remove this subject mean column
    data$subjMean <- NULL

    return(data)
}
## Summarizes data, handling within-subjects variables by removing inter-subject variability.
## It will still work if there are no within-S variables.
## Gives count, un-normed mean, normed mean (with same between-group mean),
##   standard deviation, standard error of the mean, and confidence interval.
## If there are within-subject variables, calculate adjusted values using method from Morey (2008).
##   data: a data frame.
##   measurevar: the name of a column that contains the variable to be summariezed
##   betweenvars: a vector containing names of columns that are between-subjects variables
##   withinvars: a vector containing names of columns that are within-subjects variables
##   idvar: the name of a column that identifies each subject (or matched subjects)
##   na.rm: a boolean that indicates whether to ignore NA's
##   conf.interval: the percent range of the confidence interval (default is 95%)
summarySEwithin <- function(data=NULL, measurevar, betweenvars=NULL, withinvars=NULL,
                            idvar=NULL, na.rm=FALSE, conf.interval=.95, .drop=TRUE) {

  # Ensure that the betweenvars and withinvars are factors
  factorvars <- vapply(data[, c(betweenvars, withinvars), drop=FALSE],
    FUN=is.factor, FUN.VALUE=logical(1))

  if (!all(factorvars)) {
    nonfactorvars <- names(factorvars)[!factorvars]
    message("Automatically converting the following non-factors to factors: ",
            paste(nonfactorvars, collapse = ", "))
    data[nonfactorvars] <- lapply(data[nonfactorvars], factor)
  }

  # Get the means from the un-normed data
  datac <- summarySE(data, measurevar, groupvars=c(betweenvars, withinvars),
                     na.rm=na.rm, conf.interval=conf.interval, .drop=.drop)

  # Drop all the unused columns (these will be calculated with normed data)
  datac$sd <- NULL
  datac$se <- NULL
  datac$ci <- NULL

  # Norm each subject's data
  ndata <- normDataWithin(data, idvar, measurevar, betweenvars, na.rm, .drop=.drop)

  # This is the name of the new column
  measurevar_n <- paste(measurevar, "_norm", sep="")

  # Collapse the normed data - now we can treat between and within vars the same
  ndatac <- summarySE(ndata, measurevar_n, groupvars=c(betweenvars, withinvars),
                      na.rm=na.rm, conf.interval=conf.interval, .drop=.drop)

  # Apply correction from Morey (2008) to the standard error and confidence interval
  #  Get the product of the number of conditions of within-S variables
  nWithinGroups    <- prod(vapply(ndatac[,withinvars, drop=FALSE], FUN=nlevels,
                           FUN.VALUE=numeric(1)))
  correctionFactor <- sqrt( nWithinGroups / (nWithinGroups-1) )

  # Apply the correction factor
  ndatac$sd <- ndatac$sd * correctionFactor
  ndatac$se <- ndatac$se * correctionFactor
  ndatac$ci <- ndatac$ci * correctionFactor

  # Combine the un-normed means with the normed results
  merge(datac, ndatac)
}
```